---
title: "Acuity"
subtitle: "Opinionated Quarto template inspired by Edward Tufte's principles"
author: "Ruslan Basyrov"
bibliography: bibliography.bib
abstract: |
  Acuity is a Quarto extension built on Tufte's principles: generous margins for sidenotes, an accessible color palette, and typography that gets out of the way. This document demonstrates each feature in context.
---

# Typography

Acuity sets body text in Atkinson Hyperlegible Next^[@AtkinsonHyperlegibleFont2019], a typeface designed by the Braille Institute for maximum legibility. Headings use EB Garamond for typographic contrast, and code is set in JetBrains Mono. All three load from Google Fonts.

# Layout & Margins

The page is a two-column grid: a body column for prose and a margin column for annotation. Figures, citations, and footnotes flow into the margin rather than interrupting the text — the margin speaks when the body cannot pause.


Wideblock elements span both columns when a figure or table is too wide for the body alone:

::: {.wideblock}
![A wide image spanning body and margin columns.](https://placehold.co/1100x300/88CCEE/black?text=Wide+Image){#fig-wide-test}
:::

# Code

Code blocks carry a left border instead of a background fill. The border marks the block as code without enclosing it in a box that fights the page color. Inline `code` uses a faint tint of the accent color.

```{r}
#| label: fig-sin
#| column: margin
#| fig-cap: "Margin figures can include plots. This one anchors to the code section."
plot(sin, -pi, 2*pi)
```

# Color Palette

[Paul Tol's color palette](https://sronpersonalpages.nl/~pault/), designed to remain distinguishable under color-blindness, provides the named swatches. Semantic roles — primary, success, warning, danger — are mapped separately in `_brand.yml` and drive callout colors and link accents.

```{r}
#| label: fig-palette
#| echo: false
#| message: false
#| fig-align: "center"
#| fig-cap: "Paul Tol's color palette — ten named swatches optimized for color-blind readability."

brand <- yaml::read_yaml("_extensions/acuity/_brand.yml")
pal <- unlist(brand$color$palette)
pal <- pal[!names(pal) %in% c("border")]

n_cols <- 4
df <- data.frame(
  name = names(pal),
  hex = pal,
  x = (seq_along(pal) - 1) %% n_cols,
  y = -((seq_along(pal) - 1) %/% n_cols)
)

library(ggplot2)
ggplot(df, aes(x = x, y = y, fill = hex)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(
    aes(label = sprintf("%s\n%s", name, hex)),
    color = ifelse(colSums(col2rgb(df$hex)) > 500, "black", "white"),
    fontface = "bold",
    size = 3.5
  ) +
  scale_fill_identity() +
  theme_void() +
  theme(plot.margin = margin(5, 5, 5, 5))
```

Callouts inherit semantic colors from the palette:

::: {.callout-note}
**Info** — cyan `#88CCEE`
:::

::: {.callout-tip}
**Success** — green `{{< brand color green >}}`
:::

::: {.callout-warning}
**Warning** — sand `#DDCC77`
:::

::: {.callout-important}
**Danger** — wine `#882255`
:::
